---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data

```{r, echo=TRUE}
setwd("/Users/christinabrady/Documents/Coursera courses/Data Science Specialization/Reproducible Research/RepData_PeerAssessment1")
library(ggplot2)
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
activity$interval <- as.factor(activity$interval)
```
## What is mean total number of steps taken per day?

```{r, echo=TRUE}
daily <- aggregate(steps~date, data=activity, sum)
hist(daily$steps, breaks=50)
mean(daily$steps); median(daily$steps)
```
The mean and median total number of steps taken per day are both 10766.19. This accords with what the histogram shows. 

## What is the average daily activity pattern?
```{r, echo=TRUE}
intervals <- aggregate(steps~interval, data=activity, mean)
i <- ggplot(intervals, aes(x=interval, y=steps, group=1)) + geom_line()
i
```
```{r, echo=TRUE}
intervals[which.max(intervals$steps),]
```
The interval 835 contains the most steps on average across all of the days at 206.1698 steps.

## Imputing missing values
```{r, echo=TRUE}
summary(activity)
```
There are 2304 NAs in the entire data set. Since the number of steps varies greatly over the course of a day, I will use the mean of the corresponding 5 minute interval across all days to impute the NA values. 
```{r, echo=TRUE}
colnames(intervals) <- c("interval", "imputed.steps")
ai <- merge(activity, intervals)
ai$steps[is.na(ai$steps)] <- ai$imputed.steps[is.na(ai$steps)]
ai <- ai[order(ai[,3], ai[,1]),]
head(ai)
```

Imputing the missing values based on the mean number of steps taken in each interval does not change the mean number of steps taken in one day. It does change the median slightly, from 10765 to 10766.19. 
```{r, echo=TRUE}
daily2 <- aggregate(steps~date, data=ai, sum)
hist(daily2$steps, breaks=50)
mean(daily2$steps); median(daily2$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
The activity patterns for weekdays and weekends do differ. Mean activity increases rapidly on weekdays
```{r, echo=TRUE}
ai$day <- weekdays(as.Date(ai$date))
day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
day.type <- c(rep("Weekday", 5), rep("Weekend", 2))
d <- data.frame(day=day, day.type=day.type)
ai2 <- merge(ai, d)
ai2$day.type <- as.factor(ai2$day.type)
ai3 <- aggregate(steps~day.type+interval, data=ai2, mean)
dplot <- ggplot(ai3, aes(x=interval, y=steps, group=1)) +geom_line() + facet_grid(day.type~.) + xlab("Interval: \n 00 (left) to 2355 (right)")
dplot
```
On the weekdays, on average, activity peaks at 8:35 am. 
```{r, echo=TRUE}
ai3$interval[ai3$steps == max(ai3$steps[ai3$day.type=="Weekday"])]
```
Whereas on the weekends, activity peaks at 9:15am. 
```{r, echo=TRUE}
ai3$interval[ai3$steps == max(ai3$steps[ai3$day.type=="Weekend"])]
```
Furthermore, the peak is greater on weekdays, and activity levels out with minor peaks and low points throughout the day. Whereas, on weekends, the initial peak is lower, but activity remains constantly higher throughout the day. 